---
title: "2. Get Walkable areas"
author: "Esteban Moro"
date: '2022-04-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(tigris)
```

# Introduction 

In this script we use Open Street maps to find all the walkable areas in a metropolitan area. In this case we are going to do it in the Boston Metro area

# Get the metro area for Boston

First we load the core based statistica area around Boston
```{r,results = "hide", message = FALSE, warning = FALSE}
require(tigris)
require(sf)
require(tidyverse)
cbsas <- core_based_statistical_areas(class="sf",)
boston_metro <- cbsas[cbsas$GEOID=="14460",] %>%
  st_transform(crs=4326)
```

Let's see it
```{r}
ggplot(boston_metro) + geom_sf()
```



# Get the secondary and residential roads

Using Open Street Map get the secondary roads. We select only the secondary, local and park areas. Warning: the files downloaded from OSM are large. We might want to save the for future use.

```{r,eval=F}
require(osmdata)
highway_secondary <- st_bbox(map) %>%
    opq(timeout=20000)%>%
    add_osm_feature(key = "highway", 
                    value = c("secondary", "tertiary",
                              "secondary_link","tertiary_link")) %>%
    osmdata_sf()
```

We might want to save this file for future use
```{r,eval=F}
file_secondary <- "./data/14460_secondary.RData"
save(highway_secondary,file=file_secondary)
```

```{r load_seconday,echo=F}
load("./data/14460_secondary.RData")
```

Now we get the local streets
```{r,eval=F}
highway_local <- st_bbox(map) %>%
  opq(timeout=20000)%>%
  add_osm_feature(key = "highway", 
                  value = c("residential", "living_street",
                            #"service","unclassified",
                            "cycleway",
                            "pedestrian", "footway","track","path"
                  )
  ) %>%
  osmdata_sf()
```

We might want to save this file for future use
```{r,eval=F}
file_local <- "./data/14460_local.RData"
save(highway_local,file=file_local)
```

```{r load_local,echo=F}
load("./data/14460_local.RData")
```

And finally we get the polygons for the parks, where walks are also possible.
```{r,eval=F}
parks <-  st_bbox(map) %>% 
  opq(timeout=20000) %>%
  add_osm_feature(key = "leisure",
                  value = c("park","playground","golf_course",
                            "nature_reserve")
  )%>% 
  osmdata_sf()
```
We might want to save this file for future use
```{r,eval=F}
file_parks <- "./data/14460_parks.RData"
save(parks,file_parks)
```

```{r load_parks,echo=F}
load("./data/14460_parks.RData")
```


# Get the walkable areas by county subdivision

We now calculate a buffer around those streets, paths and parks of 20 meters to get the polygon within which walks are possible. 
Given that the calculations are complex and big, we split the buffer calculation by census county subdivision.

Get the county subdivisions within the Boston Metro area (they are in the states of MA and NH)
```{r,message=F,warning=F,results=F}
cts <- map_dfr(c("MA","NH"),~{
  county_subdivisions(.x, cb = TRUE, year = 2020)
  }) %>%
  st_transform(crs=4326) %>%
  st_intersection(boston_metro)
```

Now we get the buffer by census tract
```{r,message=F,warning=F,results=F}
for(i in 1:nrow(cts)){

  #select the polygon 
  cts_selected <- cts[i,]

  #buffer a little bit around it (500 meters)
  cts_selected <- st_buffer(cts_selected,dist=500*1e-5)
  
  #intersect streets and parks with county boundaries
  cts_hs <- st_intersection(highway_secondary$osm_lines,cts_selected)
  cts_hl <- st_intersection(highway_local$osm_lines,cts_selected)
  cts_pk <- st_intersection(parks$osm_polygons,cts_selected)
  
  #get 20 meters buffers around them
  buffer_hs <- st_buffer(cts_hs,dist=20*1e-5)
  buffer_hl <- st_buffer(cts_hl,dist=20*1e-5)
  buffer_pk <- st_buffer(cts_pk,dist=20*1e-5)

  #join all the buffers into one polygon
  buffer_total <- rbind(buffer_hs[,c("osm_id","name","geometry")],
                        buffer_hl[,c("osm_id","name","geometry")],
                        buffer_pk[,c("osm_id","name","geometry")])
  buffer <- st_union(buffer_total)
  
  #remove small holes in the total buffer polygon 
  require(nngeo)
  buffer_simplified <- nngeo::st_remove_holes(buffer,max_area=(200*1e-5)^2)
  
  #save the buffer for future use
  fileOut <- paste0("./data/buffer_14460_",cts_selected$GEOID,".RData")
  save(buffer_simplified,file=fileOut)
}
```

