---
title: "3. Get walks within buffers"
author: "Esteban Moro"
date: '2022-04-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(Rcpp)
library(zoo)
library(anytime)
library(stringr)
```


# Introduction 

In this script we check whether the walks happen in the walkable buffer areas and get only those that have the majority of points within those areas. 

# Load data 

First we load the walks detected

```{r}
require(data.table)
require(tidyverse)
require(sf)
walks <- fread("./data/walks.csv.gz")
coords_walk <- st_as_sf(walks,coords=c("lon","lat"),crs=4326)
```

Get the county subdivisions within the Boston Metro area (they are in the states of MA and NH)
```{r,message=F,warning=F,results=F}
require(tigris)
cbsas <- core_based_statistical_areas(class="sf")
boston_metro <- cbsas[cbsas$GEOID=="14460",] %>%
  st_transform(crs=4326)
cts <- map_dfr(c("MA","NH"),~{
  county_subdivisions(.x, cb = TRUE, year = 2020)
  }) %>%
  st_transform(crs=4326) %>%
  st_intersection(boston_metro)
```

# Check walks within buffer

We go over the differen county subdivisions buffers (calculated in `2_Get_walkable_areas.Rmd`) to check if the points in them are within the walkable buffers.
```{r,message=F,warning=F}
walks_new <- walks
walks_new$buffer <- 0
for(i in 1:nrow(cts)){
  cts_selected <- cts[i,]
  file_buffer <- paste0("./data/buffer_14460_",cts_selected$GEOID,".RData")
  load(file_buffer)
  within_id <- st_within(coords_walk,buffer_simplified)
  ii <- which(as.numeric(within_id)==1)
  if(length(ii)>0) walks_new$buffer[ii] <- 1
}
```

# Select only walks with more the 75% points in walkable areas

```{r}
walks_new[,fraction:=sum(buffer)/.N,.(walkNum)]
walks_new <- walks_new[fraction > .75]
```

Save final set of walks
```{r}
fwrite(walks_new,file="./data/walks_final.csv.gz")
```


